Practical guides for implementing 'Five Pillar' principles

### 2023-06-07 guides/practical_guides.Rmd
added more practical guides

</details>
</p>
Source: https://github.com/markziemann/5pillars/blob/main/manuscript/5pillars.Rmd

## Practical guides

@@

### Getting started with data analysis in R and Python

@@

* Intro to R and Rstudio

* Intro to Python and iPython IDE

* Intro to vscode

### Literate programming with R Markdown, Jupyter and Quarto

@@

* https://medium.com/@itsjzt/beginner-guide-to-markdown-229adce30074

* https://www.dataquest.io/blog/r-markdown-guide-cheatsheet/

* https://tysonbarrett.com/Rstats/chapter-9-reproducible-workflow-with-rmarkdown.html

* https://andrewbtran.github.io/NICAR/2017/reproducibility-rmarkdown/rmarkdown.html

* https://quarto.org/docs/computations/r.html

* https://quarto.org/docs/computations/python.html

* https://jupyter-notebook-beginner-guide.readthedocs.io/en/latest/

* https://www.dataquest.io/blog/jupyter-notebook-tutorial/

* https://medium.com/velotio-perspectives/the-ultimate-beginners-guide-to-jupyter-notebooks-6b00846ed2af

### Practical guides for `git` and GitHub

@@

* https://www.datacamp.com/tutorial/github-and-git-tutorial-for-beginners/

* https://www.freecodecamp.org/news/introduction-to-git-and-github/

* https://www.geo.uzh.ch/microsite/reproducible_research/post/rr-rstudio-git/

* https://anderfernandez.com/en/blog/how-to-link-and-use-git-with-rstudio/

* https://andgarc.github.io/basic-git-jupyter-lesson/index.html

* https://blog.reviewnb.com/github-jupyter-notebook/

* https://andgarc.github.io/basic-git-jupyter-lesson/index.html

### Practical guides to data sharing

@@

### Extend the scripts to make them end-to-end processes

@@

### Practical guides for Conda, Guix and Docker

@@

### Practical guides for documenting computational research

@@

### Moving from "dev" to "prod": testing and continuous analysis

@@

>>>>

</details>
</p>

Source: https://github.com/markziemann/5pillars/blob/main/manuscript/practical_guides.Rmd

@@

## Getting started with data analysis in R and Python

@@

### Intro to R and Rstudio

* [Ebook: R for data science](https://r4ds.had.co.nz/)

* [Video: Rstudio for the Total Beginner](https://www.youtube.com/watch?v=FIrsOBy5k58)

* [Ebook: YaRrr! The Pirate’s Guide to R](https://bookdown.org/ndphillips/YaRrr/)

### Shell and linux systems

* [Ebook: Bioinformatics Data Skills](https://vincebuffalo.com/book/)

* [Book: Unix Shell Programming (ISBN:0672324903)](https://www.amazon.com/Unix-Shell-Programming-Stephen-Kochan/dp/0672324903)

### Intro to Python

* [Tutorial: Beginner's Guide to Python](https://wiki.python.org/moin/BeginnersGuide)

* [Ebook: The Absolute Beginner's Guide to Python Programming](https://doi.org/10.1007/978-1-4842-8716-3)

* [Video: Python Tutorial for Absolute Beginners](https://www.youtube.com/watch?v=Z1Yd7upQsXY&t=19s)

### Intro to VS Code

* [Tutorial and video: Getting Started with Visual Studio Code](https://code.visualstudio.com/docs/introvideos/basics)

## Literate programming with R Markdown, Jupyter and Quarto

@@

* [Tutorial: Beginner Guide to Markdown](https://medium.com/@itsjzt/beginner-guide-to-markdown-229adce30074)

* [Tutorial: Getting Started with R Markdown — Guide and Cheatsheet](https://www.dataquest.io/blog/r-markdown-guide-cheatsheet/)

* [Ebook Chapter: Reproducible Workflow with R Markdown](https://tysonbarrett.com/Rstats/chapter-9-reproducible-workflow-with-rmarkdown.html)

* [Tutorial: Reproducibility and R Markdown](https://andrewbtran.github.io/NICAR/2017/reproducibility-rmarkdown/rmarkdown.html)

* [Quick guide: R based literate scripts with Quarto](https://quarto.org/docs/computations/r.html)

* [Quick guide: Python based literate scripts with Quarto](https://quarto.org/docs/computations/python.html)

* [Tutorial: Jupyter/IPython Notebook Quick Start Guide](https://jupyter-notebook-beginner-guide.readthedocs.io/en/latest/)

* [Tutorial: How to Use Jupyter Notebook](https://www.dataquest.io/blog/jupyter-notebook-tutorial/)

* [Tutorial: The Ultimate Beginner’s Guide to Jupyter Notebooks](https://medium.com/velotio-perspectives/the-ultimate-beginners-guide-to-jupyter-notebooks-6b00846ed2af)

## Practical guides for `git` and GitHub

@@

* [Tutorial: GitHub and Git for Beginners](https://www.datacamp.com/tutorial/github-and-git-tutorial-for-beginners/)

* [Tutorial: How to Use Git and GitHub – Introduction for Beginners](https://www.freecodecamp.org/news/introduction-to-git-and-github/)

* [Tutorial: How to use Git with R and RStudio](https://www.geo.uzh.ch/microsite/reproducible_research/post/rr-rstudio-git/)

* [Tutorial: How to link and use Git with RStudio](https://anderfernandez.com/en/blog/how-to-link-and-use-git-with-rstudio/)

* [Tutorial: Introduction to Using Git in Jupyter](https://andgarc.github.io/basic-git-jupyter-lesson/index.html)

* [Tutorial: How to use Git / GitHub with Jupyter Notebook](https://blog.reviewnb.com/github-jupyter-notebook/)

* [Tutorial: Introduction to Using Git in Jupyter](https://andgarc.github.io/basic-git-jupyter-lesson/index.html)

## Practical guides to data sharing

@@

* [Guide: Dryad good data practices](https://datadryad.org/stash/best_practices)

* [Guide: Best practice for managing your outputs on Figshare](https://help.figshare.com/article/best-practice-for-managing-your-outputs-on-figshare)

* [Guide: Introduction to data sharing](https://zenodo.org/badge/DOI/10.5281/zenodo.3518692.svg)

* [Guide: FAIR Data (according to Australian Research Data Commons)](https://ardc.edu.au/resource/fair-data/)

## Extend the scripts to make them end-to-end processes

@@

* [Ebook: The Practice of Reproducible Research - Case Studies and Lessons from the Data-Intensive Sciences](http://www.practicereproducibleresearch.org/)

* [Ebook: The Turing Way - Guide for Reproducible Research](https://the-turing-way.netlify.app/reproducible-research/reproducible-research.html#)

* [Ebook: Building reproducible analytical pipelines with R](https://raps-with-r.dev/)

* [Tutorial: A short introduction to drake (An R based workflow manager)](https://data.nozav.org/posts/2019-a-short-introduction-to-drake/)

* [Tutorial: An example Snakemake workflow](https://snakemake.readthedocs.io/en/stable/tutorial/basics.html)

* [Tutorial compendium: Learn Nextflow in 2023](https://snakemake.readthedocs.io/en/stable/tutorial/basics.html)

## Practical guides for Conda, Guix and Docker

@@

* [Tutorial: Getting started with conda](https://conda.io/projects/conda/en/latest/user-guide/getting-started.html)

* [Video tutorial: Anaconda Beginners Guide for Linux and Windows](https://www.youtube.com/watch?v=MUZtVEDKXsk)

* [Tutorial: Running RStudio Server with Docker](https://davetang.org/muse/2021/04/24/running-rstudio-server-with-docker/)

* [Tutorial: Creating a Docker Image for your Application](https://www.stereolabs.com/docs/docker/creating-your-image/)

* [Documentation: Getting started with GNU Guix](https://guix.gnu.org/manual/en/html_node/Getting-Started.html)

## Practical guides for documenting computational research

@@

* [Guide: How to write an Awesome README](https://towardsdatascience.com/how-to-write-an-awesome-readme-68bf4be91f8b)

* [Example: README for "A Reproducible Data Analysis Workflow with R Markdown, Git, Make, and Docker"](https://github.com/aaronpeikert/reproducible-research/blob/master/README.md)

## Moving from "dev" to "prod": testing and continuous analysis

@@

* [Ebook: Building reproducible analytical pipelines with R](https://raps-with-r.dev/)

* [Ebook: The Turing Way - Guide for Reproducible Research](https://the-turing-way.netlify.app/reproducible-research/reproducible-research.html#)

* [Tutorial: Continuous Analysis](https://greenelab.github.io/continuous_analysis/)
